---
title: "Next-generation sequencing bulk segregant analysis with QTLseqr"
author: "Ben N. Mansfeld and Rebecca Grumet"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
abstract: > 
  Originally developed in the early 1990’s, Bulk Segregant Analysis (BSA) has since been an extremely useful tool for rapidly identifying markers in a genomic region associated with a trait of interest.QTL-seq is a relatively new and rapid way for performing bulk segregant analysis using next-generation sequencing data. However, no easy-to-use and multi-system compatible algorithms for performing this analysis are readily available.We developed the QTLseqr R package to allow for quick identification of genomic regions associated with traits of interest by combining two methods for bulk segregant analysis, $G'$ and $\Delta SNP-index$, described by Magwene et al. (2011) and Tagaki et al. (2013), respectively.Plotting of data is implemented in ggplot2 which allows for high-quality and customizable figures. QTLseqr package  version `r packageVersion("QTLseqr")` 
vignette: >
  %\VignetteIndexEntry{Next-generation sequencing bulk segregant analysis with QTLseqr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Standard workflow

**If you use QTLseqr in published research, please cite:**

> Mansfeld B.N. and Grumet R,
> QTLseqr: An R package for bulk segregant analysis with next-generation sequencing
> *Journal* 2017, **other info**

## Quick Start
Here are the basic steps required to run and plot QTLseq and G' analysis

```{r quickStart, eval=FALSE}

#load the package
library("QTLseqr")

#Set sample and file names
HighBulk <- "SRR834931"
LowBulk <- "SRR834927"
file <- "SNPs_from_GATK.table"
#Choose which chromosomes will be included in the analysis (i.e. exclude smaller contigs)
Chroms <- paste0(rep("Chr", 12), 1:12)

#Import SNP data from file
df <-
    importFromGATK(
        filename = file,
        highBulk = HighBulk,
        lowBulk = LowBulk,
        chromList = Chroms
     )

#Filter SNPs based on some criteria
df_filt <-
    filterSNPs(
        df,
        refAlleleFreq = 0.20,
        minTotalDepth = 100,
        maxTotalDepth = 400,
        minSampleDepth = 40,
        minGQ = 99
    )


#Run G' analysis
df_filt <- runGprimeAnalysis(
    df_filt,
    windowSize = 1e6,
    outlierFilter = "deltaSNP")

#Plot
plotQTLStats(df_filt, var = "deltaSNP", plotThreshold = TRUE, q = 0.01)
plotQTLStats(df_filt, var = "Gprime", plotThreshold = TRUE, q = 0.01)
```

## Input data

QTLseqr currently only supports table format SNP data exported from the VariantsToTable function built in to GATK. We hope to support import from SAMtools/BCFtools aswell as a simple table format, soon.

### Import from GATK
We highly recommend reading [What is a VCF and how should I interpret it?](http://gatkforums.broadinstitute.org/gatk/discussion/1268/what-is-a-vcf-and-how-should-i-interpret-it) for more information on GATK Fields and Genotype Fields

Though the use of GATK's VariantsToTable function is out of the scope of this vignette, the syntax for use with QTLseqr should look something like this:

```
java -jar GenomeAnalysisTK.jar \
-T VariantsToTable \
-R ${REF} \
-V ${NAME} \
-F CHROM -F POS -F REF -F ALT \
-GF AD -GF DP -GF GQ -GF PL \
-o ${NAME}.table
```
Where `${REF}` is the reference genome file and `${NAME}` is VCF file you wish to parse.

To run QTLseqr succesfully, the required fields `(-F)` are CHROM (Chromosome) and POS (Position). the required Genotype fields `(-GF)` are AD (Allele Depth), DP (Depth), GQ (Genotype Quality). Recommended fields are REF (Reference allele) and ALT (Alternative allele) Recommended Genotype fields are PL (Phred-scaled likelihoods)





The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
